---
title: "Web_scraping"
output: pdf_document
date: "2025-04-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Section A
in this section firstly we try to Retrieve the data from the summary table „Forecast for the next 48 hours“ and summarize its information. Firstly we read the whole of HTML file named"Weather for Sari, Iran.html". The city I have choosen is Sari, a city in Iran. In the code below I have find the table with html_elements and looking for a text with the value of ""Forecast for the next 48 hours"". This section has lots of table but the one we are looking for is the first one so I used position() function and make it equal to 1. So the main table is assigned to main table and can be seen below.
This table consists of the 2 days, Monday and Tuesday, and the times of the day. 


```{r}

library("rvest")

main_html = read_html("Weather for Sari, Iran.html")

selected_section <- main_html %>% html_elements(xpath = '//*[text() = "Forecast for the next 48 hours"]') %>% html_elements(xpath = '..')
main_table <- selected_section %>% html_elements(xpath = './/table[position() = 1]')
information = main_table %>% html_table()
print(information)
```



## Section B
In this section I visualize the predicted development of temperature, wind, and humidity for the next 24 hours (retrieve the data from „Detailed Hourly Forecast - Next 24 hours“), firstly I take the html file named "Hourly forecast for Sari, Iran.html". and get the related rows. I useed lapply and rbind to combine the rows together. Below you can see all of the information in related to wether in 24 hours in this city.

```{r}

hourly_data_html <- read_html('Hourly forecast for Sari, Iran.html')
selected_section_hourly = hourly_data_html %>% html_elements(xpath = '//*[@id= "wt-hbh-title"]') %>% html_elements(xpath = '..')
main_hourly_rows <- selected_section_hourly %>% html_elements(xpath = './/table[position() = 1]//tr[position() > 1]')


table <- main_hourly_rows %>%
  lapply(function(row) {
    row %>% html_elements("td") %>% html_text()
  }) %>% do.call(rbind, .) %>% as.data.frame()

colnames(table) <- c( "Conditions" ,"Temp" , 	"Weather",	"Feels" , "Like" ,	"Wind" ,	 	"Humidity" ,	"Chance" , "Amount")

print(table)

```
As you can see in the table above all of the nubers are charachter and they are not the pure number. For example the have some strings like : "°C". we have to eliminate them and get the number. In order to this I used "gsub()" function which eliminate the other strings and return just the numbers. We have to use function "as.numeric()" to convert it to numeric.
Now we can use this numbers to illustrate the scatter plot.

```{r}

Humidity <- gsub("%", "", table[1 :23 , "Humidity"])
Humidity <- as.numeric(Humidity)

Temp <- gsub("[^0-9]", "", table[1 :23 , "Temp"])
Temp <- as.numeric(Temp)

Feels <- gsub("[^0-9]", "", table[1 :23 , "Feels"])
Feels <- as.numeric(Feels)

print(Humidity)
print(Temp)
print(Feels)




```
 Here we firstly use these numbers to a dataframe. and than use ggplot and geom_point() to draw them.

```{r}
library(ggplot2)

Humidity_df <- data.frame(
  Hour = 1:23,                    # X-axis: hours 1 to 23
  Humidity_col = Humidity[1:23]   # Y-axis: humidity values
)

Temp_df <- data.frame(
  Hour = 1:23,                    # X-axis: hours 1 to 23
  Temp_col = Temp[1:23]           # Y-axis: humidity values
)

Feels_df <- data.frame(
  Hour = 1:23,                    # X-axis: hours 1 to 23
  Feels_col = Feels[1:23]         # Y-axis: humidity values
)

ggplot(Humidity_df,  aes(x = Hour , y = Humidity_col)) + geom_point(size = 2) + labs(x = "Hour", y = "Humidity", title = "Humidity  Over Hours") 

ggplot(Temp_df,  aes(x = Hour , y =  Temp_col)) + geom_point(size = 2) + labs(x = "Hour", y = "Temp", title = "Temp  Over Hours") 


ggplot(Feels_df,  aes(x = Hour , y = Feels_col)) + geom_point(size = 2) + labs(x = "Hour", y = "Feels", title = "Feels  Over Hours") 

```




```{r}

annual_temp_avarage = read_html("Climate & Weather Averages in Sari, Iran.html")
information <- annual_temp_avarage %>% html_elements(xpath = '//*[@id= "climateTable"]//p') %>% html_text()
print(information)


```


```{r}
information_high_temp <- annual_temp_avarage %>% html_elements(xpath = '//*[@id= "climateTable" ]//div[@class = "four columns" and position()=1]//p[position() = 1]') %>% html_text()

information_low_temp <- annual_temp_avarage %>% html_elements(xpath = '//*[@id= "climateTable" ]//div[@class = "four columns" and position()=1]//p[position() = 2]') %>% html_text()

information_mean_temp <- annual_temp_avarage %>% html_elements(xpath = '//*[@id= "climateTable" ]//div[@class = "four columns" and position()=1]//p[position() = 3]') %>% html_text()
information_precipitation <- annual_temp_avarage %>% html_elements(xpath = '//*[@id= "climateTable" ]//div[@class = "four columns" and position()=2]//p[position() = 1]') %>% html_text()

print(information_high_temp)
print(information_low_temp)
print(information_mean_temp)
print(information_precipitation)

```

```{r}
high_temp <- as.numeric(gsub("[^0-9]", "", information_high_temp))
low_temp <- as.numeric(gsub("[^0-9]", "", information_low_temp))
mean_temp <- as.numeric(gsub("[^0-9]", "", information_mean_temp))
precipitation <- as.numeric(gsub("[^0-9]", "", information_precipitation))/10


print(high_temp)
print(low_temp)
print(mean_temp)
print(precipitation)

```


```{r}

final_table <- rbind(high_temp , low_temp, mean_temp, precipitation)
colnames(final_table) <- c("All year" , "January" , "February" , "March" , "April" , "May", "June" , "July" , "August" , "September" , "October" , "November" , "December")
print(final_table)

```

```{r}

High_df <- data.frame(
  Month = 2:13,                    
  High = high_temp[2:13]           
)

Low_df <- data.frame(
  Month = 2:13,                     
  Low = low_temp[2:13]          
)

Mean_df <- data.frame(
  Month = 2:13,                    
  Mean = mean_temp[2:13]          
)

Precision_df <- data.frame(
  Month = 2:13,                   
  Precision = precipitation[2:13]          
)




ggplot(High_df,  aes(x =Month , y = High )) + geom_point(size = 2) + labs(x = "Months", y = "High Avarage" ) 

ggplot(Low_df, aes(x = Month , y = Low )) + geom_point(size = 2) + labs(x = "Months", y = "Low Avarage" ) 

ggplot(Mean_df , aes(x = Month , y = Mean )) + geom_point(size = 2) + labs(x = "Months", y = "Mean" ) 


ggplot(Precision_df , aes(x = Month , y = Precision )) + geom_point(size = 2) + labs(x = "Months", y = "Precision" ) 



```


```{r}
annual_max <- high_temp[1]
annual_min <- low_temp[1]
annual_mean <- mean_temp[1]
annual_precision <- precipitation[1]

print(paste("The annual min tempreture is :" , annual_max, "°C"))
print(paste("The annual max tempreture is :" , annual_min , "°C"))
print(paste("The annual mean tempreture is :" , annual_mean, "°C"))
print(paste("The annual precision  is :" , annual_precision, "mm"))

```







